## Geocode address and map to census tract

This example starts with a street address and then displays two maps:

 - state and county of interest,
 - census tracts within county of interest.

Useful background info about 2000 Census, but most is applicable to 2010: [US Census Spatial and Demographic Data in R](http://www.r-project.org/conferences/useR-2010/slides/Almquist.pdf).

Earl F Glynn, UMKC Center for Health Insights, 2014-10-27. 

### Setup
```{r setup}
library(stringr)   # str_trim
library(RJSONIO)   # fromJSON

fix <- function(s)
{
  gsub(" ", "+", str_trim(s))
}

# http://stackoverflow.com/questions/6364783/capitalize-the-first-letter-of-both-words-in-a-two-word-string
simpleCap <- function(x)
{
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
```

### Test Address:  UMKC School of Medicine

```{r address}
street <- "2411 Holmes St"
city   <- "Kansas City"
state  <- "MO"
```

### Output from US Census Geocoder

Form URL for [US Census geocoder](# http://geocoding.geo.census.gov/geocoder/) and parse JSON output:
```{r geocoder}
GEOCODE.PREFIX <- "http://geocoding.geo.census.gov/geocoder/geographies/address?street="
GEOCODE.SUFFIX <- "&benchmark=Public_AR_Census2010&vintage=Census2010_Census2010&layers=14&format=json"

URL <- paste0(GEOCODE.PREFIX, fix(street), "&city=", fix(city), "&state=", fix(state), GEOCODE.SUFFIX)

JSON <- fromJSON(URL)

length(JSON$result$addressMatches)
stopifnot(length(JSON$result$addressMatches) > 0)
```

Values of interest from geocoder:
```{r values1}
target.matched <- JSON$result$addressMatches[[1]]$matchedAddress
target.matched

target.stateabbr <- JSON$result$addressMatches[[1]]$addressComponents["state"]
target.stateabbr

target.zip <- JSON$result$addressMatches[[1]]$addressComponents["zip"]
target.zip

target.coordinates <- JSON$result$addressMatches[[1]]$coordinates
target.coordinates
```

It's unclear why geographies object is a vector here.  Make sure there is a least one, but defer decision on how to handle mutiple geographics till later. 

See [Wikipedia info about FIPS codes](# http://en.wikipedia.org/wiki/Federal_Information_Processing_Standard_state_code#FIPS_state_codes).

```{r values2}
length(JSON$result$addressMatches[[1]]$geographies$`Census Blocks`)
stopifnot(length(JSON$result$addressMatches[[1]]$geographies$`Census Blocks`) > 0)

fips.state   <- JSON$result$addressMatches[[1]]$geographies$`Census Blocks`[[1]]$STATE
fips.state

fips.county  <- JSON$result$addressMatches[[1]]$geographies$`Census Blocks`[[1]]$COUNTY
fips.county

target.tract <- JSON$result$addressMatches[[1]]$geographies$`Census Blocks`[[1]]$TRACT
target.tract  # US Census tract

target.block <- JSON$result$addressMatches[[1]]$geographies$`Census Blocks`[[1]]$BLOCK
target.block  # US Census block

target.geoid <- JSON$result$addressMatches[[1]]$geographies$`Census Blocks`[[1]]$GEOID
target.geoid
```

### State and County Map

```{r statemap}
library(maps)
data(state.fips)
data(county.fips)

fips.state.name <- as.character(state.fips$polyname[ which(as.numeric(fips.state) == state.fips$fips)[1] ])
fips.county.name <- as.character(county.fips$polyname[ county.fips$fips == as.integer(substr(target.geoid,1,5)) ])

# Some states come in parts.  Here only state name wanted.
fips.state.name <- unlist(strsplit(fips.state.name, ":"))[1]
fips.state.name

# State
county.info <- map("county", fips.state.name, plot=FALSE)
map("county", fips.state.name,
   fill=TRUE,
   col=ifelse(county.info$names == fips.county.name, "blue", "transparent"))

map("state", fips.state.name, lwd=2, add=TRUE)
title(simpleCap(fips.state.name), cex.main=1.5)

SubdivisionName <- "County"
if (fips.state.name == "alaska")    SubdivsionName <- "Borough"
if (fips.state.name == "louisiana") SubdivsionName <- "Parish"
county.name <- unlist(strsplit(fips.county.name, ","))[2]          # "jackson"
county.name.full <- paste(simpleCap(county.name),SubdivisionName)  # "Jackson County"
mtext(county.name.full, col="blue")

points(target.coordinates["x"], target.coordinates["y"], col="green", pch=8)
```

### Census Tract Map for County

```{r countymap1}
library(UScensus2010)
library(UScensus2010tract)
```

load creates R object, e.g., missouri.tract10, from data read from file.
```{r loadtracts}
load(paste0(find.package("UScensus2010tract"), "/data/", fips.state.name, ".tract10.rda"))
```

Assign object (e.g., missouri.tract10) to standard name (state.tract10) for manipulation and delete original name.  Because of R's "lazy loading" this is effectively a rename and does not cause a duplicate copy in memory.
```{r parseeval}
R.statement <- paste0("state.tract10 <- ", fips.state.name, ".tract10;\n",
                      "rm(", fips.state.name, ".tract10)")
connection <- textConnection(R.statement)
R.parse <- parse(connection)
close(connection)
eval(R.parse)
```

### Census Tract Map for County

```{r countymap2}
county.coordinates <- county(name=county.name, state=target.stateabbr, level="tract")
plot(county.coordinates, main=county.name.full)
title(paste0(county.name.full, ", ", target.stateabbr), col.main="blue", cex.main=1.5)

plot(state.tract10,
     col=ifelse(state.tract10@data$tract == target.tract, "red", "transparent"),
     border="transparent",
     add=TRUE)
rm(state.tract10)

tract.parts <- paste(substr(target.geoid,1,2), substr(target.geoid,3,5), substr(target.geoid,6,11))
mtext(paste("Census Tract",tract.parts), col="red")

points(target.coordinates["x"], target.coordinates["y"], col="green", pch=8)
```

